A Dependently Typed Multi-Stage Calculus \ifnotanonymous\thanks{Supported by organization x.}\fi

### 2019-06-18 text/properties.tex
Minor revision.

Now, we turn our attention to proving Progress for staged semantics.
The following lemma Unique Decomposition ensures that, every typed
term is either a value or can be uniquely decomposed to an evaluation
context and a redex, ensuring that a well typed term is not
immediately stuck and the staged semantics is deterministic.

\begin{theorem}[Unique Decomposition]
  If $\G$ does not have any variable declared at stage $\varepsilon$ 
  and $\G \V M : \tau @ A$, then either
  \begin{enumerate}
  \item $ M \in V^A$, or
  \item there uniquely exist $B, E^A_B$, and $R^B$ such that $M = E^A_B[R^B]$.
  \end{enumerate}
\end{theorem}

\begin{proof}
  We can prove by straightforward induction on typing derivations.
  Main cases are \TApp, \TTBL, and \TIns.
  We need Inversion Lemmas for them.
\end{proof}

In a usual setting, the type environment $\G$ has to be empty, namely
the term has to be closed.  The condition is relaxed here because
variables at stages higher than \(\varepsilon\) are considered
symbols.  In fact, this relaxation is required for proof by induction
to work.

@@

\begin{corollary}[Progress]
	If $\G$ does not have any variable declared at stage $\varepsilon$ and $\G \V M : \tau  @ A$ then
	$ M \in V^A $ or $M'$ exists such that $M \longrightarrow_s M'$.
\end{corollary}

>>>>

Now, we turn our attention to staged semantics.  First, the staged
reduction relation is a subrelation of full reduction, so Subject
Reduction holds also for the staged reduction.

\begin{theorem}
  If $M \longrightarrow_s M'$, then $M \longrightarrow M'$.
\end{theorem}
\begin{proof}
  Easy.
 \qed
\end{proof}

The following theorem Unique Decomposition ensures that, every typed
term is either a value or can be uniquely decomposed to an evaluation
context and a redex, ensuring that a well typed term is not
immediately stuck and the staged semantics is deterministic.

\begin{theorem}[Unique Decomposition]
  If $\G$ does not have any variable declared at stage $\varepsilon$ 
  and $\G \V M : \tau @ A$, then either
  \begin{enumerate}
  \item $ M \in V^A$, or
  \item $M$ can be uniquely decomposed into an evaluation context and a redex, that is, there uniquely exist $B, E^A_B$, and $R^B$ such that $M = E^A_B[R^B]$.
  \end{enumerate}
\end{theorem}

\begin{proof}
  We can prove by straightforward induction on typing derivations.
  Main cases are \TApp, \TTBL, and \TIns, where we use Inversion.
  \qed
\end{proof}

The type environment $\G$ in the statement usually has to be empty;
in other words, the term has to be closed.  The condition is relaxed here
because variables at stages higher than \(\varepsilon\) are considered
symbols.  In fact, this relaxation is required for proof by induction
to work.

@@

\begin{corollary}[Progress]
	If $\G$ does not have any variable declared at stage $\varepsilon$ and $\G \V M : \tau  @ A$, then
	$ M \in V^A $ or $M'$ exists such that $M \longrightarrow_s M'$.
\end{corollary}
